### 2015-01-17 docs/report/title.tex
Added initial docs files

\begin{titlepage}
\begin{center}

\textsc{\LARGE University of Buenos Aires}\\[1.5cm]

\textsc{\Large Professional Assignment}\\[0.5cm]

% Title
{ \huge \bfseries Using algorithmic skeletons in EcmaScript to parallelize computation in browsers through Web Workers \\[0.4cm] }

% Author and supervisor
\noindent

\emph{Authors:}\\
Damian \textsc{Schenkelman}\\
Matias \textsc{Servetto}\\[0.5cm]

\emph{Tutor:}\\
Prof.~Rosa \textsc{Wachenhauzer}

\vfill

% Bottom of the page
{\large \today}

\end{center}
\end{titlepage}