### 2015-02-08 docs/report/code_sharing.tex
Completed code_sharing.tex

\section{Code sharing}

One of the aspects that must be handled by the library is the distribution of the code between the different workers when a skeleton is created.

A naive approach would be trying to send the function to be executed directly to the worker. The following code fragment attempts to do so:
\lstinputlisting[caption=Trying to send a function to a worker, escapechar=]{../../spikes/2.2.1/test-functionSerialization/funcSerialization.js}

When executed in the Google Chrome console the error shown in Figure~\ref{fig:func-serialization} is thrown.
\img{Error when trying to send a Function object to a Web Worker}{../../spikes/2.2.1/test-functionSerialization/funcSerialization.png}{fig:func-serialization}

That is an expected error as explained in section 9.5.3 of the HTML standard \cite{html-whatwg}, the value being sent must support the structured clone algorithm and a Function does not meet that requirement as explained in section 2.7.5 of that same standard. It could be assumed that this is because of the challenges of serializing function closures\cite{cloud-haskell}\cite{distributed-scheme}.

\subsection{Serializing functions}
Considering the fact that \tfunction{} objects cannot be directly transferred a different serialization approach is to be considered. Since function closures will not be transferrable a possible alternative is to:
\begin{enumerate}
  \item Serialize the function to a \tstring{}.
  \item Transfer the \tstring{} to the worker.
  \item Create a new function on the worker from that \tstring{}.
\end{enumerate}

\note{An approach to provide contextual data to worker functions is explored in \todo{link to section}.}

\subsubsection{\tstring{} serialization benchmarks}
One possible way of passing serialized \tfunction{} to workers would be to encode them as binary and transfer them as an \tabuffer{}. To encode the strings there are two possible approaches:
\begin{itemize}
  \item Using the Encoding API\cite{encoding-api}
  \item Implementing non native encoding/decoding functions
\end{itemize}

We created a benchmark to understand the difference between each approach:
\bmcode{http://jsperf.com/pjs-serialization-comparison/2}{../../spikes/2.2.1/test-syncEncodingAPI/encodingAPI.js}

The results for it are displayed in Figure~\ref{fig:string-encode-decode}.
\img{Comparing string encoding/decoding approaches}{../../spikes/2.2.1/test-syncEncodingAPI/encodingAPI}{fig:string-encode-decode}

\subsubsection{Conclusion}
The native implementation was faster than the non-native one and the difference increased between Chrome versions, which seems to indicate that some extra optimizations could be expected in the future.

\subsection{Transferring strings}
Function's code not only needs to be serialized but also transferred. For that reason it was also worth comparing the time it takes to encode/decode a \tstring{} and tranfer the resulting \tabuffer{} to the worker against just passing the \tstring{} to the worker.

Additionally, in most scenarios code would have to be sent along with a \ttarray{} so it would also be interesting to see if the transfer time was affected by this fact.

\subsubsection{\tstring{} serialization and tranfer benchmarks}
We created a benchmark using two functions of different length to see if the function's length affected the serialization and transfer time. We measured three approaches to transfer the string:
\begin{enumerate}
  \item Encoding the \tstring{} into a \tabuffer{} and sending it as a \code{Tranferable} and decoding it on the Web Worker.
  \item Sending the string directly
  \item Using a \code{Blob} to store the stream and retrieving that in the Worker.
\end{enumerate}

The benchmark with the short function was the following one:
\bmcode{http://jsperf.com/pjs-serialization/2}{../../spikes/2.2.1/test-howToSendCode/shortSerialization.js}

The results for it are displayed in Figure~\ref{fig:short-serialization}.
\img{Comparing string encoding/decoding approaches}{../../spikes/2.2.1/test-howToSendCode/shortSerialization}{fig:short-serialization}

The benchmark with the long function was the following one:
\bmcode{http://jsperf.com/pjs-serialization-long}{../../spikes/2.2.1/test-howToSendCode/longSerialization.js}

The results for it are displayed in Figure~\ref{fig:long-serialization}.
\img{Comparing string encoding/decoding approaches}{../../spikes/2.2.1/test-howToSendCode/longSerialization}{fig:long-serialization}

\subsubsection{\tstring{} and \ttarray{} serialization and tranfer benchmarks}
We created a benchmark that tranfers a \ttarray{} and \tstring{} back and forth between the main thread an a worker to understand if transferring additional objects affected the transfer time when comparing it with just tranferring a \tstring{}.

The benchmark's code is the following one:
\bmcode{http://jsperf.com/pjs-encoding}{../../spikes/2.2.1/test-howToSendCodeWithData/encoding.js}

The results for it are displayed in Figure~\ref{fig:code-with-data}.
\img{Comparing string encoding/decoding approaches}{../../spikes/2.2.1/test-howToSendCodeWithData/encoding}{fig:code-with-data}

\subsubsection{Conclusion}
As the benchmarks show, the fastest way to tranfer the \tstring{} between the UI thread and workers is by directly sending it.

\pagebreak