### 2015-02-14 docs/report/initialization.tex
Fixes in report

\section{Initialization}

As explained in the Web Workers standard \cite{w3c-ww} \textit{``Generally, workers are expected to be long-lived, have a high start-up performance cost, and a high per-instance memory cost."}.

There are two factors to consider from the above sentence regarding the library's implementation:

\begin{description}
\item[First] The high start-up time.
\item[Second] The high per-instance memory cost.
\end{description}

\subsection{High start-up time}
This factor raises the issue of when workers should be instantiated. On relation to this, we performed a benchmark to get an idea of the difference between cold starting a worker against having an already available one process messages.

\subsubsection{Cold start benchmark}
The following is the code for the benchmark:
\bmcode{http://jsperf.com/worker-cold-start/5}{../../spikes/2.1.1/newVsPreExistent.js}

The results for it are displayed in Figure~\ref{fig:new-vs-pre}.
\img{Evaluating the inefficiency of cold starting a Web Worker}{../../spikes/2.1.1/newVsPreExistent}{fig:new-vs-pre}

\subsubsection{Conclusion}
The previous benchmark shows a huge difference between the time required to process a message on a pre-existent Web Worker and that necessary to start Web Worker and process a message. For that reason, we decided to create a pool of Web Workers when the library is initialized.

\subsection{High per-instance memory cost}
The second factor makes it important for users to be able to control/limit the amount of web workers that the library instantiates. For that reason the library allows users to specify a maximum amount of workers to be instantiated, and also provide through its API a way to terminate all workers in existance.

\subsection{Maximum amount of workers}
Since the work being done is CPU intensive (i.e.: the workers' goal is not to perform I/O operations), the maximum amount of workers to be created should not be greater than the amount of cores in the machine \(W_{cores}\), otherwise interleaving would start to affect performance. Additionally, considering the necessity of users to control how many workers are created, an upper bound can be specified \(W_{upper bound}\). If no upper bound is specified, the user upper bound for workers is assumed to be infinite (because the user is not providing a restriction). To summarize, the number of workers to be instantiated is:
\[W = min(W_{cores}, W_{upper bound})\]

\pagebreak