### 2015-05-24 docs/report/function_transfer_time.tex

\section{Decreasing function transfer time}\label{sec:function-transfer-time}

Each time \code{seq} is invoked on a chain, all callback functions related to \code{map}, \code{filter} and \code{reduce} steps are serialized and sent to the Web Workers.

Each Web Worker has a cache to avoid using different instances of the same function (as explained in Section~\ref{sec:func-caching}). Nevertheless, if the same function is going to be used multiple times in a program as the callback for steps, it might make sense to send it to the Web Workers once and then simply make a reference to it.

This can be achieved by taking advantage of the global context object (introduced in Section~\ref{sec:context}).

The first thing that needs to be done is to call \code{pjs.updateContext} by providing a \tfunction{} as a property of the context object. An example is shown in the following code snippet:
\begin{lstlisting}[caption=Sending function in global context]
var promise = pjs.updateContext({
  add: function (x) { return x + 2; },
});
\end{lstlisting}

Once the global context has been updated, the \tstring{} representation of the function's key (\code{'add'}) can be used as a callback parameter instead of using a \tfunction{} instance. An example is shown in the following code snippet:
\begin{lstlisting}[caption=Passing a function name from the global context instead of a \tfunction{} as a callback]
var promise = pjs.updateContext({
  add: function (x) { return x + 2; },
}).then(function(){
  pjs(new Uint8Array([1,2,3,4])).map('add').seq()
    .then(function(result){
      // result is [ 3, 4, 5, 6 ]
    });
});
\end{lstlisting}

\subsection{Benchmark}
We created a benchmark to analyze the improvement obtained by just sending a \tstring{} instead of a \tfunction{} instance.

The following listing shows the code for the benchmark:
\bmcode{http://jsperf.com/global-ctx-inlined-func/3}{../../spikes/6.2/inlinedVsGlobalInlined/jspBenchmark.js}

The results for the benchmark are displayed in Figure~\ref{fig:func-key-vs-instance}.
\img{Comparing sending function key in global context against \tfunction{} instance}{../../spikes/6.2/inlinedVsGlobalInlined/result}{fig:func-key-vs-instance}

\subsection{Conclusion}
While the performance improvement obtained is not remarkable, it is still noticeable and does not require a lot of effort from a developer's standpoint. It is worth pointing out that the improvement shown in Figure~\ref{fig:func-key-vs-instance} will vary if the difference in character length between the key and the function's code does.

\pagebreak