### 2015-05-25 docs/report/array_merging.tex

\section{Array merging}

Once all the workers have executed the \tfunction{} provided to them on each element of their portion of the original array, they are to notify the UI thread and provide the resulting \tabuffer{} or \tsabuffer{}.

At this point the UI thread must consolidate the results from the Web Workers into a new \ttarray{} or \tstarray{} that represents the result for the entire operation.\footnote{The case for chains with \code{reduce} as the last step similar and simpler. Thus, it is not worth analyzing.}

For the \tabuffer{} case, this means copying all the elements of each partial result into a larger \ttarray{} while maintaining the original order.

For the \tsabuffer{} case there are two possible situations:
\begin{enumerate}
  \item If all steps in the chain perform the \code{map} transformation, then no changes are required.
  \item If there is at least one \code{filter} step, the elements need to be placed at the beginning of the resulting \tstarray{} in order to avoid \textit{holes}.
\end{enumerate}

The latter is similar to the \tabuffer{} case, but instead of copying from one array to another, the source and target are the same.

\subsection{\ttarray{} merge benchmark}
Unlike the case for \ttarray{} splitting (Section~\ref{subsec:copying-typed-arrays}) there aren't that many native alternatives to create a \ttarray{} from smaller ones.

The benchmark presented only has one alternative that uses a method with a native implementation:
\bmcode{http://jsperf.com/typedarray-merge/2}{../../spikes/2.4.1/merge.js}

The results for the benchmark are displayed in Figure~\ref{fig:typed-array-merge}.
\img{Comparing \ttarray{} merge approaches}{../../spikes/2.4.1/merge}{fig:typed-array-merge}

\subsection{\tstarray{} merge benchmark}

The benchmark modifies the \ttarray{} merge benchmark to support instances of the \tstarray{} type. The implementation does not include the \dataview{} alternative as it is not supported by \tstarray{}s:
\bmcode{http://jsperf.com/sharedtypedarray-merge}{../../spikes/2.4.1/sharedMerge.js}

The results for the benchmark are displayed in Figure~\ref{fig:shared-typed-array-merge}.
\img{Comparing \tstarray{} merge approaches}{../../spikes/2.4.1/sharedMerge}{fig:shared-typed-array-merge}

\subsection{Conclusion}
The benchmark shows that the best alternative to merge small arrays into a larger one is using \code{Set} method of either \ttarray{} or \tstarray{}.

For illustration purposes, a naive implementation (without optimizations or error checking) would be similar to the following one:
\begin{lstlisting}[caption=Simple \ttarray{} merge function]
function merge(arrays){
  var first = arrays[0];
  var total = arrays.reduce(function(c,a) { return c + a.length; }, 0);
  var result = new first.constructor(total);
  var start = 0;

  arrays.forEach(function(a){
    result.set(array, start);
    start += array.length;
  });

  return result;
}
\end{lstlisting}

\pagebreak